### 2018-04-18 paper.Rmd
made comments
addressed writing comments through line 246

author: 
  - name          : "Alessandro Sanchez*"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, 450 Serra Mall, Stanford, CA 94305"
    email         : "sanchez7@stanford.edu"
  - name          : "Stephan Meylan*"
    affiliation   : "2"
  - name          : "Mika Braginsky"
    affiliation   : "3"
  - name          : "Kyle E. MacDonald"
    affiliation   : "1"
  - name          : "Daniel Yurovsky"
    affiliation   : "4"
  - name          : "Michael C. Frank"
    affiliation   : "1"

@@

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development.
  Access to these data can be complex for novices and difficult to automate for advanced users, however.
  To address these issues, we introduce *childes-db*, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API).
  Along with a versioned infrastructure that facilitates reproducibility of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

Children's language acquisition is a core puzzle in cognitive science.
How are the representations underlying the productive use of language learned over children's first years?
Because a language is comprised of many individual pieces of information -- the names of particular objects, the marker for the past tense, the position of determiners -- language development is fundamentally a question of learning from environmental input, which is the only source of information on these specifics. <!-- SM: innate knowledge? generalzation? -->
Further, children are often reluctant experimental participants; often their most revealing and most sophisticated uses of language emerge in naturalistic play. <!-- SM: we could motivate this differently-- it's not just that experiments are hard with young children; it's that many phenomena can only be seen in naturalistic data -->
Thus, language acquisition research must be grounded in the ecological behavior of children, both their language input and their language production in context.

The foundational resource for studies of language acquisition is recordings and transcripts of children. <!-- SM: this needs to be qualified--- Piaget, Berko, etc. experimental and diary studies predate naturalistic trancscripts -->
Starting with Roger Brown's [-@brown1973] work on Adam, Eve, and Sarah, longitudinal recordings of children's talk (and the talk they hear from their caregivers) have formed the basis for this field. <!-- SM: what field? child language? see above comment -->
But creating individual transcripts is extremely hard work.
Beyond gathering the data, even orthographic transcriptions (without phonetic detail) can take ten times the length of the original recording to transcribe [@macwhinney2000].<!--SM: qualify with "relatively simple" orthographic transcriptions; may need to define `phonetic` for general psych or (non-language) child dev audience -->
Thus, researchers' desire for data from multiple ages or multiple children can very quickly outstrip their resources.

Established in 1984 to address this issue, the Child Language Data Exchange System (CHILDES) aims to make transcripts and recording of children's language available for researchers via a free resource [@macwhinney2000].<!--SM: we might want to cite the whole set of CHILDES releases, starting with the earliest one in 1984 -->
CHILDES archives tens of thousands of transcripts and related media, making it a critical resource for characterizing children’s early productive language use as well as their language environment. <!-- SM: added a point about related media b/c Brian MacWhinney thinks this is a central point of CHILDES; tens of thousands of transcripts of children's speech +[across n languages] -->
As the first major effort to consolidate and share transcripts of child language, CHILDES has been a pioneer in the move to curate and disseminate large-scale behavioral datasets publicly.

Since its inception, a tremendous body of research has made use of CHILDES data. <!-- SM: switched corpus -> body to avoid confusion -->
Individual studies are too numerous to list, but classics include studies of morphological over-regularization [@marcus1992], distributional learning [@redington1998], word segmentation [@goldwater2009], the role of frequency in word learning [@goodman2008] and many others. <!-- SM: to convey its importance, we could say how many citations CHILDES main papers have received and how many citations papers that use CHILDES have received -->
Some studies analyze individual examples in depth [e.g., @snyder2007], other track multiple child-caregiver dyads [e.g., @meylan2017], and still others use the aggregate properties of child or caregiver speech. <!--SM: need a citation for using the aggregate speech --> 
These are just a few of the thousands of studies that take CHILDES corpora -- transcripts, annotations, and associated media -- as starting points.

Yet there are some challenges working with CHILDES, both for students and for advanced users.
The CHILDES ecosystem typically uses a special file format (CHAT), which is stored as plain text but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological, morphological, or syntactic structure of the utterance. <!-- added "typically" b/c BMW will note the availability of XML -->
These files are typically analyzed using a command-line tool (CLAN) with a host of specific arguments that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU), and execture complex searches against the data.
While this system is flexible and powerful, mastering the CHAT codes and especially the CLAN tool can be daunting for a beginner or student.
These technical barriers decrease the ease of casual exploration by a novice researcher or in a classroom exercise.

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal.
CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- it can be tricky to build into a processing pipeline written in a data analysis platform such as the Python Pandas pacakage or R. <!--analytic -> data analysis platform -->
Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analyses typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

The practice of writing custom parsers is problematic for a number of reasons.
First, effort is wasted in implementing the same features again and again.
Second, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard.
Third, these parsing scripts are rarely shared, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017 for an example].
Fourth, the CHILDES corpus itself is a moving target; computational work using the entire corpus at one time point may include a different set of data than subsequent work (due to newly-contributed corpora or revised annotations of existing ones) and there is no simple way for researchers to document exactly what version of the corpus has been used short of creating a full mirror of the data.
These factors together lead to a lack of computational reproducibility, a major problem that keeps researchers from verifying or building on published research [@donoho2010;@stodden2016].

In the current manuscript, we describe a system for extending the functionality of CHILDES to address these issues.
Our system, childes-db, is a database-formatted mirror of CHILDES that allows programmatic access through an application programming interface (API). 
This infrastructure allows the creation of web applications for browsing and visualizing the data easily, facilitating classroom use of the dataset. 
Further, it can be accessed programmatically by advanced researchers, obviating the need to write one-off parsers of the CHAT format.
The database is versioned for access to previous releases, allowing computational reproducibility of particular analyses.

We begin by describing the architecture of childes-db and the web applications that we provide.
We next describe the `childesr` API, which allows for programmatic access to the data.
We then present several worked examples of specific uses of the system for research and teaching using both the web app and the API. 
<!-- a little too much "we" in this final paragraph; needs a final hopefil statement -->

@@

<!-- Remind readers why they're reading this: understanding the structure helps you use and understand the tools. SM: Moved this up from database format because it addresses more general issues -->
As described above, CHILDES is most often approached as a set of distinct CHAT files, which are then parsed by users (often using CLAN). 
In contrast to this parsing approach, which entails the sequential processing of strings, childes-db treats CHILDES as a set of linked tables, with records corresponding to intuitive abstractions such as words, utterances, and transcripts.
Users of data analysis languages like R or Julia, libraries like Pandas, or those familiar with Structured Query Language (SQL) will be familiar with operations on tables such as filtering (subsetting), sorting, aggregation (grouping), and joins (merges). 
These operations obviate the need for users to consider the specifics of the CHAT representation -- instead they simply request the entities they need for their research and allow the API to take care of the formatting details. 
We begin by orienting readers to the design of the system via a top-level description and motivation for the design of the database schema, details of the database’s current technical implementation, and our versioning scheme. 
Users primarily interested in accessing the database can skip these details and focus on access through the `childesr` API and the web apps. 
<!-- We first present the design and technical approach of childes-db, then describe the interactive analysis tools provided by our web-based frontend.  -->

@@

Other tables in the database are hierarchical collections built out of tokens -- *utterance*, *transcript*, *corpus*, and *collection* -- which store attributes appropriate for each level of description.
Every entity includes attributes that link it to all higher-order collections, e.g., an utterance lists the transcript, corpus, and collection to which it belongs.
An *utterance* contains one or more words and includes fields such as the utterance type (declarative, question, etc.), total number of tokens, and the total number of morphemes (if the morphological structure is available in the original CHAT).
A *transcript* consists of one or more utterances and includes the date collected, the name of the target child, and age in days (if defined), and the filename from CHILDES.
A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus.
Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES.
Because every record can be linked to a top-level collection (generally corresponding to a language), each table includes data from all languages represented in CHILDES.

Participants (e.g., children) are represented separately from the token hierarchy because it is common for the same children to appear in multiple transcripts.
A participant identifier is associated with every word and utterance, including a name, role, 3-letter CHILDES identifier (CHI = child, MOT = mother, FAT = father, etc.), and the range of ages (or age of corresponding child) for which they are observed.
For non-child participants (caregivers and others), the record additionally contains an identifier for the corresponding target child to facilitate associating caregivers with children.

@@

childes-db is implemented as a MySQL database, an industry-standard, open-source relational database server that can be accessed directly from a wide range of programming languages.
The childes-db project provides hosted read-only databases for direct access and for `childesr` (described below) as well as compressed .sql exports for local installation.
While the former is appropriate for most users, local installation can provide performance gains (allowing a user to access the database on their local machine or on their network) as well as allowing users to store derived information in the same database.

@@

The content of CHILDES changes as additional corpora are added or transcriptions are updated; as of the time of writing, these changes are not systematically tracked.
We introduce a simple versioning system by adding a new complete parse of the current state of CHILDES every six months or as warranted by changes in CHILDES.
By default, users interact with the most recent version of the database available. 
But to support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last three years / six versions) through our `childesr` API so that researchers can run analyses against specific historical versions of the database.
For versions more than three years old, we host compressed .sql files that users may download and serve using a local installation of MySQL server.
<!-- SM: How easy is it to implement versioning in the Shiny apps? This would show anyone using the web app the availability of the versions AS: app functionality depends on db version in some cases, so nontrivial SM: let's punt-->

@@

The current implementation of childes-db emphasizes the computation of lexical statistics, and consequently focuses on reproducing the words, utterances, and speaker information in CHILDES.
We have yet to parse and make available the following information:

@@

Our web apps are built using Shiny, an R package that enables easy app construction using `R`.
Underneath the hood, each web app is making calls to our `childesr` API and subsequently plotting the data using the popular R plotting package `ggplot2`.
A user's only task is to configure exactly what should be plotted through a series of buttons, sliders, and text boxes.
The user may specify what collection, corpus, child, age range, caregiver, etc., should be included in a given analysis.
The plot is displayed and updated in real-time, and the underlying data are also available for download alongside the plot to facilitate customization. All of these analyses may also be reproduced using the `childesr` package.

@@

We have currently implemented:

- MLU-w (mean length of utterance in words, a common measure of complexity),^[Because of gaps in available CHILDES morphological data, MLU-w is computed and not MLU-m (mean length of utterance in morphemes). 
<!-- sm: I think we should go ahead and compute this for transcript with MOR tiers? Given that this was one of the first uses, I don't think Brian will like the 'underlying gaps' explanation. MCF: I agree. -->
In English, these two measures are highly correlated. <!-- sm: citation needed -->
alongside those of children.] 
- TTR (type-token ratio, a measure of lexical diversity), 
- MTLD (measure of textual lexical diversity), 
- HD-D (lexical diversity via the hypergeometric distribution), 
- and voc-d (a further measure of lexical diversity).  <!-- sm: MLU is not a measure of lexical diversity-->
<!-- MCF: need citations for voc-d HD-dand MTLD -->

@@

For example, an instructor could create a programming assignment with the goal of reproducing the key findings in the case studies presented above -- color words or gender. 
epending on the students' knowledge of R, the instructor could decide how much of the `childesr` starter code to provide before asking students to generate their own plots and write-ups.
Moreover, the instructor could take advantage of recent advances in tools for teaching literate programming such as R Markdown [@xie2015dynamic], building in pre-written blocks of executable code that demonstrate key programming concepts for students and including incomplete or blank code blocks that students can use to write their own code. <!-- SM: cool, but scope creep? -->

>>>>

author: 
  - name          : "Alessandro Sanchez*"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, 450 Serra Mall, Stanford, CA 94305"
    email         : "sanchez7@stanford.edu"
  - name          : "Stephan C. Meylan*"
    affiliation   : "2"
  - name          : "Mika Braginsky"
    affiliation   : "3"
  - name          : "Kyle E. MacDonald"
    affiliation   : "1"
  - name          : "Daniel Yurovsky"
    affiliation   : "4"
  - name          : "Michael C. Frank"
    affiliation   : "1"

@@

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development, particularly in characterizing the early language learning environment.
  Access to these data can be complex for novices and difficult to automate for advanced users, however.
  To address these issues, we introduce *childes-db*, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API).
  Along with a versioned infrastructure that facilitates reproducibility of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

How children acquire language is one of the core puzzles in cognitive science, developmental psychology, and linguistics. 
What are the representation that children learn about language, and how to they emerge from the interaction of learning mechanisms and environmental input?
Developing facility with language requires learning a great many interlocking components -- meaningful distinctions between sounds (phonology), names of particular objects and actions (word learning), meaningful sub-word structure (morphology), rules for how to organize words together (syntax), and context-dependent and context-independent aspects of meaning (semantics and pragmatics).
Key to learning all of these systems is the contribution of the child's input -- exposure to linguistic and non-linguistic data -- in the early environment.
While in-lab experiments can shed light on linguistic knowledge and some of the implicated learning mechanisms, characterizing this early environment requires additional research methods and resources.

One of the key methods that has emerged to address this gap is the collection and annotation of speech to and by children, often in the context of the home. 
Starting with Roger Brown's [-@brown1973] work on Adam, Eve, and Sarah, audio recordings -- and more recently augmented with video recordings -- have been augmented with rich, searchable annotations that have allowed researchers to address a number of questions regarding the language learning environment.
Focusing on language learning in naturalistic contexts also reveals that children have in many cases productive and receptive abilities exceeding those revealed in experimental contexts; owing to challenges in lab-based experiments with young children, often their most revealing and sophisticated uses of language emerge in naturalistic play.

While corpora of early language acquisition are extremely useful, creating such corpora requires significant resources. 
Collecting and transcribing audio and video is costly and extremely time consuming  -- even orthographic transcription (i.e., transcriptions with minimal phonetic detail) can take ten times the length of the original recording to transcribe [@macwhinney2000].
Automated, machine learning -based methods like automatic speech recognition (ASR) have provided only modest gains in efficiency, in that they are limited by the less-than-ideal acoustic properties of home recordings and the poor fit of language models built on adult-directed, adult-produced language samples for characterizing child-directed or child-produced speech.
Thus, researchers' desire for sufficient sample sizes and coverage in analyses of child language corpora can very quickly outstrip their resources. 
<!-- data pooling is necessary) -->

Established in 1984 to address this issue, the Child Language Data Exchange System (CHILDES) aims to make transcripts and recording of children's language available for researchers as free, public datasets [@macwhinney1985; @macwhinney2000; @macwhinney2014].
CHILDES archives tens of thousands of transcripts and associated media across 20+ languages, making it a critical resource for characterizing children’s early productive language use as well as their language environment. 
As the first major effort to consolidate and share transcripts of child language, CHILDES has been a pioneer in the move to curate and disseminate large-scale behavioral datasets publicly.

Since its inception, a tremendous body of research has made use of CHILDES data.
Individual studies are too numerous to list, but classics include studies of morphological over-regularization [@marcus1992], distributional learning [@redington1998], word segmentation [@goldwater2009], the role of frequency in word learning [@goodman2008], and many others.
Some studies analyze individual examples in depth [e.g., @snyder2007], other track multiple child-caregiver dyads [e.g., @meylan2017], and still others use the aggregate properties of child or caregiver speech [e.g., @redingtonEtAl1998; @montagEtAl2015].
These are just a few of the thousands of studies that make use of CHILDES corpora.

Nonetheless, there are some outstanding challenges working with CHILDES, both for students and for advanced users.
The CHILDES ecosystem typically uses a specialized file format (CHAT), which is stored as plain text but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological, morphological, or syntactic structure of the utterance.
These files are usually analyzed using a command-line tool (CLAN), which has a number of specific tools that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU), and execute complex searches against the data.
While this system is flexible and powerful, mastering the CHAT codes and especially the CLAN tool with its many tools and flags can be daunting for a beginner or student.
These technical barriers decrease the ease of exploration by a novice researcher or in a classroom exercise.

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal.
CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- it can be tricky to build into a processing pipeline written in Python or R.
Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analyses typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

The practice of writing custom parsers is problematic for a number of reasons.
First, effort is wasted in implementing the same features again and again.
Second, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard.
Third, these parsing scripts are rarely shared or break with subsequent revisions to the dataset, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017 for an example].
Fourth, the CHILDES corpus itself is a moving target; computational work using the entire corpus at one time point may include a different set of data than subsequent work due to newly-contributed corpora or revised annotations of existing ones; currently, there is no simple way for researchers to document exactly what version of the corpus has been used short of creating a full mirror of the data.
These factors together lead to a lack of computational reproducibility, a major problem that keeps researchers from verifying or building on published research [@donoho2010;@stodden2016].

In the current manuscript, we describe a system for extending the functionality of CHILDES to address these issues.
Our system, childes-db, is a database-formatted mirror of CHILDES that allows programmatic access through an application programming interface (API). 
This infrastructure allows the creation of web applications for browsing and easily visualizing the data, facilitating classroom use of the dataset. 
Further, it can be accessed programmatically by advanced researchers, obviating the need to write one-off parsers of the CHAT format.
The database is versioned for access to previous releases, allowing computational reproducibility of particular analyses.

We begin by describing the architecture of childes-db and the web applications that we provide.
Next, we describe the `childesr` API, which provides a set of R functions for programmatic access to the data.
We conclude by presenting several worked examples of specific uses of the system--both web apps and the R API--for both research and teaching.

@@

As described above, CHILDES is most often approached as a set of distinct CHAT files, which are then parsed by users (often using CLAN). 
In contrast to this parsing approach, which entails the sequential processing of strings, childes-db treats CHILDES as a set of linked tables, with records corresponding to intuitive abstractions such as words, utterances, and transcripts.
Users of data analysis languages like R or Julia, libraries like Pandas, or those familiar with Structured Query Language (SQL) will be familiar with operations on tables such as filtering (subsetting), sorting, aggregation (grouping), and joins (merges). 
These operations obviate the need for users to consider the specifics of the CHAT representation -- instead they simply request the entities they need for their research and allow the API to take care of the formatting details. 
We begin by orienting readers to the design of the system via a top-level description and motivation for the design of the database schema, details of the database’s current technical implementation, and our versioning scheme. 
Users primarily interested in accessing the database can skip these details and focus on access through the `childesr` API and the web apps. 
<!-- SM: not sure if we want to have readers skip this; there may be some conceptual dependencies here -->
<!-- We first present the design and technical approach of childes-db, then describe the interactive analysis tools provided by our web-based frontend.  -->

@@

Manty of the other tables in the database are hierarchical collections built out of tokens -- *utterance*, *transcript*, *corpus*, and *collection* -- which store attributes appropriate for each level of description.
Every entity includes attributes that link it to all higher-order collections, e.g., an utterance lists the transcript, corpus, and collection to which it belongs.
An *utterance* contains one or more words and includes fields such as the utterance type (declarative, question, etc.), total number of tokens, and the total number of morphemes (if the morphological structure is available in the original CHAT).
A *transcript* consists of one or more utterances and includes the date collected, the name of the target child, and age in days (if defined), and the filename from CHILDES.
A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus.
Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES.
Because every record can be linked to a top-level collection (generally corresponding to a language), each table includes data from all languages represented in CHILDES.

Participants (e.g., children and caregivers) are represented separately from the token hierarchy because it is common for the same children to appear in multiple transcripts.
A participant identifier is associated with every word and utterance, including a name, role, 3-letter CHILDES identifier (CHI = child, MOT = mother, FAT = father, etc.), and the range of ages (or age of corresponding child) for which they are observed.
For non-child participants (caregivers and others), the record additionally contains an identifier for the corresponding target child, such that data correpsonding to children and their caregivers can be easily associated.

@@

childes-db is implemented as a MySQL database, an industry-standard, open-source relational database server that can be accessed directly from a wide range of programming languages.
The childes-db project provides hosted, read-only databases for direct access and for `childesr` (described below) as well as compressed .sql exports for local installation.
While the former is appropriate for most users, local installation can provide performance gains (allowing a user to access the database on their machine or on their local network) as well as allowing users to store derived information in the same database.

@@

The content of CHILDES changes as additional corpora are added or transcriptions are updated; as of the time of writing, these changes are not systematically tracked.
We introduce a simple versioning system by adding a new complete parse of the current state of CHILDES every six months or as warranted by changes in CHILDES.
By default, users interact with the most recent version of the database available. 
To support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last three years / six versions) through our `childesr` API so that researchers can run analyses against specific historical versions of the database.
For versions more than three years old, we host compressed .sql files that users may download and serve using a local installation of MySQL server.
<!-- SM: How easy is it to implement versioning in the Shiny apps? This would show anyone using the web app the availability of the versions AS: app functionality depends on db version in some cases, so nontrivial SM: let's punt-->

@@

The current implementation of childes-db emphasizes the computation of lexical statistics, and consequently focuses on reproducing the words, utterances, and speaker information in CHILDES transcripts.
We have yet to parse and make available the following information:

@@

Our web apps are built using Shiny, an R package that enables easy app construction using `R`.
Underneath the hood, each web app is making calls to our `childesr` API and subsequently plots the data using the popular R plotting package `ggplot2`.
A user's only task is to configure exactly what should be plotted through a series of buttons, sliders, and text boxes.
The user may specify what collection, corpus, child, age range, caregiver, etc., should be included in a given analysis.
The plot is displayed and updated in real-time, and the underlying data are also available for download alongside the plot to facilitate customization. All of these analyses may also be reproduced using the `childesr` package.

@@

We have currently implemented the following measures:

- MLU-w (mean length of utterance in words, a common measure of complexity),^[Because of gaps in available CHILDES morphological data, MLU-w is computed and not MLU-m (mean length of utterance in morphemes). 
<!-- sm: I think we should go ahead and compute this for transcript with MOR tiers? Given that this was one of the first uses, I don't think Brian will like the 'underlying gaps' explanation. MCF: I agree. -->
In English, these two measures are highly correlated. <!-- sm: citation needed -->
alongside those of children.] 
- TTR [type-token ratio, a measure of lexical diversity; @templin1957], 
- MTLD (measure of textual lexical diversity), 
- HD-D (lexical diversity via the hypergeometric distribution), 
- and voc-d (a further measure of lexical diversity).  <!-- sm: MLU is not a measure of lexical diversity-->
<!-- MCF: need citations for voc-d HD-dand MTLD -->

@@

For example, an instructor could create a programming assignment with the goal of reproducing the key findings in the case studies presented above -- color words or gender. 
Depending on the students' knowledge of R, the instructor could decide how much of the `childesr` starter code to provide before asking students to generate their own plots and write-ups.
Moreover, the instructor could take advantage of recent advances in tools for teaching literate programming such as R Markdown [@xie2015dynamic], building in pre-written blocks of executable code that demonstrate key programming concepts for students and including incomplete or blank code blocks that students can use to write their own code. <!-- SM: cool, but scope creep? -->
